---
title: "Inequalities by ethnicity in the CHD pathway"
lang: en-GB
author: "Jacqueline Grout"
date: last-modified
date-format: "YYYY-MM-DD"
format:
  html:
    grid:
      sidebar-width: 250px
      body-width: 800px
      margin-width: 250px
      gutter-width: 1.5rem
    embed-resources: true
    smooth-scroll: true
    theme: cosmo
    fontcolor: black
    toc: true
    toc-location: left
    toc-title: Contents
    toc-depth: 3
editor: visual
execute:
  echo: false
  message: false
  warning: false
  freeze: auto
editor_options: 
  chunk_output_type: console
---

```{r}
#| echo: false
 
library(targets)

library(tidyverse)
```

## Introduction

In the earlier report ..........

## CHD Metrics

jhjkhhjhkj

## Clustered GP Practices

```{r}
tar_read(cluster2_map)
```

### Cluster descriptions

#### Cluster 1 - Least diverse

This cluster is the least diverse of the five and the median percentage of White patients is 97% (94% White British).

::: panel-tabset
## Five categories

```{r}
tar_read(cluster2_treemap_1)
```

## Fourteen categories

94% White British

```{r}
tar_read(cluster2_eth_chart_1)
```
:::

#### Cluster 2

In this cluster the median percentage of White patients is 93% (87% White British). Those patients whose ethnicity is White are more likely to be White Irish or other White ethnicities (0.7% and 4% respectively) when compared to cluster 1. The median percentage of patients with a mixed ethnicity in this cluster is 2%.

::: panel-tabset
## Five categories

```{r}
tar_read(cluster2_treemap_2)
```

## Fourteen categories

87% White British

```{r}
tar_read(cluster2_eth_chart_2)
```
:::

#### Cluster 3

In this cluster the median percentage of White patients is 78% (69% White British). The median percentage of patients with a mixed ethnicity in this cluster is 3.8%, which is higher than in cluster 2. Indian patients make up 3.3% of the patients, and 2.6% are patients whose ethnicity is Black African.

::: panel-tabset
## Five categories

```{r}
tar_read(cluster2_treemap_3)
```

## Fourteen categories

69% White British

```{r}
tar_read(cluster2_eth_chart_3)
```
:::

#### Cluster 4

In cluster 4 just over half the patients are White with a median percentage of 55% (35% White British). The ethnic group of White has fewer White British than clusters 1 to 3, with 17% of patients having an ethnicity of White Other The median percentage of patients whose ethnicity is Black African in this cluster is 8.3%, which is higher than in cluster 3 and the median percentage of Black Caribbean patients is 3.9%. The median percentage of Black and Asian is 29% in this cluster.

::: panel-tabset
## Five categories

```{r}
tar_read(cluster2_treemap_4)
```

## Fourteen categories

35% White British

```{r}
tar_read(cluster2_eth_chart_4)
```
:::

#### Cluster 5 - Most diverse

This cluster is the most diverse. Overall the median percentage of Asian patients in this cluster is 45%. There are 13.3% Indian, 9.7% Pakistani and 2.9% Bangladeshi. Mixed and other ethnic groups form a large proportion of the patients in this cluster.

::: panel-tabset
## Five categories

```{r}
tar_read(cluster2_treemap_5)
```

## Fourteen categories

22% White British

```{r}
tar_read(cluster2_eth_chart_5)
```
:::

## Relative Index of Disparity

```{r}
tar_read(rel_iod_chart)
```

```{r}
tar_read(ci_iod_chart)
```

## Rate

### Risk factors

```{r}
tar_read(rate_chart_risk_fact)
```

### Risk factor identification

```{r}
tar_read(rate_chart_risk_fact_ident)
```

### Primary prevention

```{r}
tar_read(rate_chart_prim_prevent)

```

### Disease identification

```{r}
tar_read(rate_chart_disease_ident)
```

### Secondary prevention

```{r}
tar_read(rate_chart_second_prevent)
```

### Tertiary prevention

```{r}
tar_read(rate_chart_tert_prevent)
```

### Intermediate outcome

```{r}
tar_read(rate_chart_int_out)
```

### Full outcomes

```{r}
tar_read(rate_chart_full_out)
```

## Appendix

### Definitions and data sources of pathway metrics

### Methods explained

#### CHD synthetic prevalence estimates

CHD synthetic prevalence estimates data has been removed from Fingertips by OHID.

OHID have sent the archived data. This is for 7564 practices from 2015, of which 6297 are current practices.

Need to identify practices with missing data and work out how to get a prevalence for them.

Also have file from previous project - checked data is the same but need to check practices included - maybe there are more?

nearest neighbours to impute chd prevalence

#### Census 2021 Ethnicity by LSOA

#### Which date should be used for the gp list size dataset?
